```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, eval = FALSE, 
                      message = FALSE, warning = FALSE)
#xaringanExtra::use_clipboard()
```

```{r}
library(tidyr)
library(dplyr)
```

## Example 1
These counts measure the number of TB cases in given county and year.
```{r}
table4a
```

- What are the variables in this data set? 

- What are the cases?


## Example 2

On a sheet of paper (or in markdown), draw how `table4a` would look if it
had the same data grouped into three columns: country, year, and cases.


## Example 3
High school graduation rates
```{r}
grad_wide<- data.frame(school=c("Aitkin","Bemidji","Chaska"), 
                gradRate_2015=c(93.0,86.2,96.2),
                gradRate_2014=c(94.6,83.2,92.7),
                region=c("northern","northern","metro"))
grad_wide
```


#### (a) Use `pivot_longer` to tidy this data set.


#### (b) Rerun your  `pivot_longer` command but include arguments that removes "gradRate" from the `names_to` column and coverts it to a integer variable type. 


#### (c) Use `separate` to divide the `gradRate_year` column into two columns named `measure` and `year`. What data type is `year` after this is done? What does adding `convert=TRUE` to the `separate command. 


## Example 4

Use `pivot_wider` to make your data set in Example 3 back into a wider version.





## Example 5
Using the `who` data set loaded through `tidyr`. This data contains counts of new tuberculosis (TB) cases broken down by year, country, age, gender, and diagnosis method from the 2014 World Health Organization Global Tuberculosis Report.

```{r}
who
```

#### (a)  Gather the 5th through 60th columns of `who` into a pair of columns named `codes` and `n`. Include `values_drop_na = TRUE` in your `pivot` command to remove cases with missing values. 

#### (b) Select just the county, year, codes and n variables

#### (c) WHO codes are encoded as: `new_<type of TB case>_<sex><age>` where
  - Type of TB case
    - `rel` - relapse
    - `ep` - extra-pulmonary
    - `sn` - pulmonary, smear negative 
    - `sp` - pulmonary, smear positive
  - Sex
    - `m` - male 
    - `f` - female
  - Age
    - `014` - 0 to 14 years old
    - `1524` - 15 to 24 years old 
    - `2534` - 25 to 34 years old 
    - `3544` - 35 to 44 years old 
    - `4554` - 45 to 54 years old 
    - `5564` - 55 to 64 years old 
    - `65` - 65 and older

Notice that `rel` does not have a `_` separator from `new` while the other three types do have a `_` separator. There is a separate `_` between type and sex but no separator bewtween sex and age. 
```{r}
names(who)
```

Try the following `separate` command (with your data name used) to attempt to  divide the `codes` column into three variables called `new`, `type` and `sexage` by the underscore separator `_`. How does the `count` command tell you that your first attempt didn't work?
```{r, eval=FALSE}
# edit this command to work with your data
your_data %>%
  separate(
    codes, 
    into = c("new", "type", "sexage"), 
    sep="_"
    ) %>% 
  count(new, type)
```

#### (d)  Now add `fill = "left"` to the command above to fill `NA`'s into columns from the left when the split by `_` doesn't contain 3 values. Verify that this does divide `sexage` and `type` correctly. 

#### (e)  Separate the sexage column into sex and age columns. Use a numeric position for `sep` to indicate how to split this column (see `?separate`).  Add `convert = TRUE` to let R take its best guess at variable type. 

