```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, eval = TRUE, 
                      message = FALSE, warning = FALSE)
#xaringanExtra::use_clipboard()
```

```{r packageCheck, include=TRUE}
library(tidyverse)
myclass <- "Math 285"
feelings <- c("fun", "early")
my_sentence <- str_c(myclass, "is", feeling, sep = " ")
```


## Problem 1 Combining

#### (a) 
What happens when a `str_c` entry is a vector? Run the following code to answer:
```{r}
my_sentences <- str_c(myclass, "is", feelings, sep = " ")
my_sentences
```


#### (b) 
What does `str_flatten` do?
```{r}
str_flatten(my_sentences, collapse = " and ")
```

#### (c) 
How to write "Math 285 is fun and early" using `str_c`?




## Problem 2 Special characters

What will using a `\n` separator do in the command below? 

```{r}
p2 <- str_c(myclass, "is", feeling, sep = "\n")
p2
writeLines(p2)
```




## Problem 3 Length
Does `str_length` count spaces and special characters??

```{r q3}
p2
str_length(p2)
```



## Problem 4 Extracting

#### (a)
What happens with negative positions? 

```{r}
str_sub(my_sentence, start = -3, end = -1)
```

#### (b)
With a vector of positions? 

```{r}
str_sub(my_sentence, start = c(1, 6), end = c(4, 8))
```

#### (c)
Only a start value?

```{r}
str_sub(my_sentence, start = -3)
```




## Problem 5 Replace

#### (a)
What does the `str_replace_all` command below do?  
```{r}
str_replace_all(my_sentence, pattern = " ", replacement = "")
```

#### (b)
What will `str_length` of the string in (a) measure?



## Problem 6 Splitting
What happens with `str_split` when `simplify = TRUE` is added?

```{r}
str_split(my_sentences, " ")
str_split(my_sentences, " ", simplify = TRUE) 
```



## Problem 7 Splitting
What happens with `str_split_fixed` if we let `n=3` or `n=5`?

```{r }
str_split_fixed(my_sentences, " ", n = 3) 
str_split_fixed(my_sentences, " ", n = 5)  
```




## Problem 8: `stringr` with data

Let's use the tools in `stringr` for some data analysis using our favorite `babynames` data. 
```{r}
library(babynames)
```


#### (a)

Use `str_sub` to create a new column in `babynames` called `last` that contains the last letter of every name. (Hint: see 4(a) above.)
```{r}
# check with a test name:
str_sub("Arnab", <fill in>)
```

```{r}
# then use to add a new variable
babynames <- babynames %>%
  mutate(last = str_sub(name, <fill in>))
```

#### (b)

Use your `last` variable created in (a) compute the proportion of names in 2017 that ended in a vowel for each sex. 

suggested way:
- make a logical variable called `vowel` that indicates whether the last letter is one of "a", "e", "i", "o", "u", or "y". (Hint: use   `%in%`) 
- filter and group, then use `vowel` to compute the count and proportion of vowel ending names by sex in 2017



#### (c)

Use `str_detect()` and other tidyverse tools to determine how many baby names in 2017 contain the string "ono" for each sex (ignoring capitalization).



#### (d)

Use `str_subset()` to extract all of the baby names in 2017 that contain "ono" (regardless of sex). Then determine which names, if any, occurred in both sexes (hint: try `duplicated`).  One way to get started is to pull all babynames in 2017 into a character vector:

```{r}
names_2017 <- babynames %>%
  filter(year == 2017) %>%
  mutate(name = str_to_lower(name)) %>%
  pull(name) 
```