```{r, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, collapse=FALSE, 
                      message = FALSE, warning = FALSE, 
                      eval = FALSE)
xaringanExtra::use_clipboard()
```


## Example 1: Calories in Food

The number of calories in five foods is given in the data set below. 
```{r message=FALSE, warning=FALSE}
library(tidyverse)
calories <- tribble(
  ~food, ~calories, ~bacon,
  "French Fries", 607, FALSE,
  "Potato chips", 542, FALSE,
  "Bacon", 533, TRUE,
  "Pizza", 296, FALSE,
  "Chili Dog", 260, FALSE
)
```

#### (a) Create a bar chart displaying the calories by food, coloring the bars by `bacon`:

- use `col` geometry
- use `fill` for the color aesthetic
- add an informative title and labels


```{r 1a solution}
# Create the bar chart
```

#### (b) Reorder the bars by caloric content.

We can use the `fct_inorder()` command to reorder the bars within our ggplot2 code since the rows are ordered. 

```{r 1b solution}
# Reorder the bars by calories
```

Other helpful functions for reordering factors for plotting include `fct_reorder()`, `fct_infreq()`, and `fct_rev()`. These functions are part of the `forcats` package which is loaded by the `tidyverse`.


#### (c) Remove the axis labels, since they are redundant.

One approach is to set their `labs()` to `NULL`:

```{r 1c solution}
# Set the x and y labels to NULL
```

Another approach is to set the appropriate element of the `theme()` to `element_blank()`:

```{r 1c solution2}
# Remove the element from the theme
```



#### (d) Remove the legend, since we are simply calling attention to bacon. (Hint: use a `theme` option.)

We can edit the appearance of the legend using a variety of arguments within the `theme()` (see `?theme` for the full list). To remove the legend, we set `legend.position = "none"` (Other options for the  position include`"left"`, `"right"`, `"bottom"`, `"top"`, or a two-element numeric vector)

```{r 1d solution}
# Remove the legend
```

#### (e) Change the scale for fill color.

The default fill colors are not ideal, especially since we are simply drawing attention to bacon. To change the scale, we appeal to a `scale_fill_<method>()` function. A few options include:

- `scale_fill_brewer()`
- `scale_fill_grey()`
- `scale_fill_viridis_d()`
- `scale_fill_manual()`


With `scale_fill_manual()`, you can manually specify named colors for a factor in the order of the levels using the `values` argument:

```{r 1e solution1}
# Change the fill colors to "gray40" and "red2"
```

Look at the `palette` options in the `?scale_fill_brewer` help file. Which type of palette, diverging, qualitative, or sequential, is best for contrasting bacon with the other foods?

```{r 1e solution2}
# Change fill color to a brewer palette
```

What happens when you add `direction = -1` to the `brewer` layer in your previous plot?
```{r e solution3}
# Change previous plot to include direction = -1
```

Try using `scale_fill_viridis_d()` to change fill color. What is one reason for using this color palette? (check the help file!)

```{r 1e solution4}
# Change color to viridis
```

#### (f) Annotate the top of the bars with the actual calories value. 

One way to add text values to each bar is with `geom_text()`. 

```{r 1f solution}
# Add text to directly label the bars
```

