```{r, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, collapse=FALSE, 
                      message = FALSE, warning = FALSE, 
                      eval = TRUE, fig.align = 'center',
                      fig.height = 4, fig.width = 6)
xaringanExtra::use_clipboard()
```


```{r packageCheck, include=TRUE, eval = TRUE}
library(ggplot2)
library(dplyr)
```


## Historical baby name {.build}
The package `babynames` contains the data set `babynames` in tibble form. This data set gives name rates for births recorded by the Social Security Administration. 

- Over 1.9 million rows giving the names of all baby names born in any year
```{r, eval = TRUE}
#install.packages("babynames")
data("babynames", package = "babynames")
babynames  # tibble prints the first 10 rows, not the entire data frame
```
- Each row specifies a:
    + `year`
    + `sex` recorded as `F` or `M`
    + `name`
    + `n` number born with that name in a given year, by sex
    + `prop` proportion with a given name, by sex and year. 
    
E.g. 7065 females named Mary were recorded in SSA applicants born in 1880, this makes up about 7.2% of all females applicants with 1880 birth dates.


\newpage

#### 1. What does this command do? 

```{r}
filter(babynames, name %in% c("Tillie","Tilly","Mathilda","Matilda"))
```



#### 2. What does this plot show?

```{r}
babynames %>% 
  filter(name == "Tillie", sex == "F") %>% 
  ggplot(aes(x = year)) + 
  geom_line(aes(y = n, color = name)) +
  scale_x_continuous(breaks = seq(1880, 2017, by=10))
```



#### 3. Use `n_distinct` to determine how many distinct years are in the data. What are the first and last years in the data?

```{r}
# for #3
```


#### 4. There are `r nrow(filter(babynames, name == "Tillie"))` years with at least one baby named Tillie. Maybe? Why might this conclusion be false? Show how you can check.

```{r}
filter(babynames, name == "Tillie")
```



#### 5. For female records, what name had the highest popularity in one year measured by the number of babies born with that name? Give the name and year. Repeat for males. 

```{r}
#  for #5
```


#### 6. Repeat 5, but measure popularity by the proportion of babies born with that name in a given year. Do the names and years differ?

```{r}
#  for #6
```


