```{r, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, collapse=FALSE, 
                      message = FALSE, warning = FALSE, 
                      eval = FALSE, fig.align = 'center', 
                      fig.height = 4, fig.width = 6)
xaringanExtra::use_clipboard()
```


```{r packageCheck}
library(ggplot2)
library(dplyr)
```


## Historical baby name
```{r}
data("babynames", package = "babynames")
babynames  # tibble prints the first 10 rows, not the entire data frame
```
- Each row specifies a:
    + `year`
    + `sex` recorded as `F` or `M`
    + `name`
    + `n` number born with that name in a given year, by sex
    + `prop` proportion with a given name, by sex and year. 
    

## Data wrangling part 1 continued...




#### 7. What does this plot show?

```{r}
babynames %>% 
  filter(sex == "F") %>% 
  count(year, name = "year_n") %>%
  ggplot(aes(x=year, y = year_n)) + 
  geom_line() +
  scale_x_continuous(breaks = seq(1880, 2017, by=10))
```



### 8. Write the following set of commands using one piped command. 

```{r}
tillies <- filter(babynames, name=="Tillie")
summarize(group_by(tillies, sex), tillie_mean = mean(n), tillie_sd = sd(n), n_years = n())
```



```{r}
#  for #8
```

### 9. Interpret the summaries given in 8.

### 10. Which year did not have any babies named Tillie for female births?


```{r}
#  for #10
```