```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse=TRUE, prompt=FALSE, 
                      eval=TRUE, message=FALSE)
```

## Objects used in this handout:
```{r}
x <- c(8,2,1,3)
w <- c(8L,2L,1L,3L)
loans <- read.csv("https://math.carleton.edu/kstclair/data/CreditData.csv")
tweets <- read.csv("https://math.carleton.edu/kstclair/data/TrumpTweetData.csv")
x_mat2 <- cbind(x, 2*x)
x_df <- data.frame(x=x, double_x=x*2)
my_list <- list(myVec=x, myDf=x_df, myString=c("hi","bye"))
```

### Problem 1: data types

(a) What data type is `x`?
(b) What data type is `w`? 
(c) What data type is `x == 1`? 
(d) What data type is `loans$Duration.in.month`? 
(e) What data type is `c(x,NA)`?

### Problem 2: Coercion

(a) What data type is `c(x, loans$Duration.in.month)`?
(b) What data type is `c(x,"NA")`?
(c) What does `sum(c(TRUE,FALSE,TRUE,FALSE))` equal?
(d) What does `sum(x < 5)` equal?



### Problem 3: Subsetting 

(a) How can we reverse the order of entries in `x`?
(b) What does `which(x < 5)` equal?
(c) What does `sum(x[c(TRUE,FALSE,TRUE,FALSE)])` equal?
(d) What does `sum(x[x < 5])` equal?
(e) Why `dim(x_mat2[1:2,1])` return `NULL` while `dim(x_mat2[1:2,1:2])` returns a dimension?



### Problem 4: Lists

(a) Using `my_list`, show three ways to write one command that gives the 3rd entry of variable `x` in data frame `myDf`
(b) What class of object does the command `my_list[3]` return?
(c) What class of object does the command `my_list[[3]]` return?
(d) What class of object does the command `unlist(my_list)` return? Why are all the entries characters?


### Question 5: Data frames 
```{r}
str(tweets$text)
```

(a) The data set `TrumpTweetData.csv` contains data collected on about 1500 Trump tweets. The variable `text` contains the text of the selected tweets. Would it make sense to convert the variable `text` to a factor variable or should we keep it a character variable? 
(b) Use **two methods** (in base-R) to find the `text` of the 180th tweet in the Trump data.
(c) What is the class of the `attributes` of the data frame `tweets`?


